A comprehensive framework to capture the arcana of neuroimaging analysis

### 2019-03-14 manuscript/manuscript.tex
Changed references to dMRI to DWI

The utility of the Arcana framework is demonstrated by the
implementation of analysis suites for T1, T2* and diffusion weighted MRI
(dMRI) data and the application of dMRI tractogram \citep{tournier_mrtrix:_2012} and
vein masks \citep{ward_vein_2017} workflows to data collected from a
healthy subject.

@@

While basic Study classes are typically associated with single image
modality or contrast, the analysis suites implemented by them can be
integrated into multi-modal analysis by aggregating multiple Study
classes (sub-studies) in a \emph{MultiStudy} class. Analysis suites are
integrated by joining the data specification tables of the sub-studies
of a MultiStudy class. Entries in specification tables of sub-studies
are joined by mapping them to a common entry in the specification table
of the MultiStudy (Figure \ref{fig:example_multi_study}). This enables derivatives from one
sub-study (e.g. brain extracted T1-weighted anatomical) to be referenced
by workflows of other sub-studies (e.g. anatomically constrained dMRI
tractography).

@@

The DwiStudy class implements the extraction of diffusion tensor metrics,
FA and ADC, as well as whole-brain tractography using streamlines tracking from the MRtrix
toolbox \citep{tournier_improved_2010,tournier_mrtrix:_2012} (Figure \ref{fig:dmri_study}).

\begin{figure*}
	\centering
    \includegraphics[width=0.55\textwidth]{../figures/dmri_study}
  \caption{Example diffusion-weighted MRI (DWI) study, which can derive
tensor metrics, fractional anisotropy (FA) and apparent diffusion
coefficient (ADC) as well as streamlines fibre tracking. Blue boxes:
acquired (input) data (filesets or fields). Green boxes: derivatives.
Orange ovals: pipelines. Blue and green arrows: acquired and derived
inputs to pipelines, respectively. Orange arrows: outputs of pipelines.
The dMRI magnitude image is preprocessed for motion correction and EPI
distortions masked and bias corrected. From the bias corrected image two
branches of analysis can be performed using the same intermediate
derivatives: FA and ADC and/or streamlines fibre tracking.}
\label{fig:dmri_study}
\end{figure*}

@@

Figure \ref{fig:fa_adc}  displays the FA and ADC maps derived from the diffusion MRI
acquisition. The FA map shows high intensity in known white matter
tracts and low intensities in known grey matter regions. The ADC map
shows high intensity in cortical spinal fluid and low intensity through
the rest of the brain.
Figure \ref{fig:tractography} displays the global tractography derived from the dMRI
acquisition. The streamlines follow well known white matter tracts such
as the cortico-spinal, fasciculus and corpus callosum. Intermediate
derivatives derived for the FA and ADC analysis, including the
preprocessed and bias-corrected dMRI image and a whole brain mask, were
reused in the generation of the streamlines.

@@

MR contrast-specific analyses are implemented in Banana via a chain
of successively specialised Study sub-classes (e.g.
MRI\textgreater{}EPI\textgreater{}dMRI) to enable generic processing
steps (e.g. registration) to be shared between classes. While not
necessary, it is recommended to create a subclass specific to the
research study in question and aggregate all related analysis within it,
since such classes can be applied to alternate datasets in order to
reproduce the exact analysis. The \emph{ArcanaPaper} class
(Supplementary material), which contains methods to generate all figures
in the Results section of this manuscript, is an example of this approach.

@@

Arcana's management of intermediate derivatives and provenance
guarantees that the same analysis is applied across the dataset without
necessarily requiring a complete rerun of the analysis. This guarantee
makes it feasible to process data as it is acquired over the course of
long studies, and therefore help detect any problems that might arise
with the acquisition protocol when they occur. In addition, by reusing
shared intermediate derivatives between analyses, such as the
preprocessed dMRI shared between tensor and fibre tracking workflows
(Figure \ref{fig:fa_adc} and \ref{fig:tractography}), processing time as
well as time required for manual
QC is minimised. Given analyses implemented in Arcana are also able to
be processed on HPC clusters, they scale well to large studies.

>>>>

The utility of the Arcana framework is demonstrated by the
implementation of analysis suites for T1, T2* and diffusion weighted MRI
(DWI) data and the application of DWI tractogram \citep{tournier_mrtrix:_2012} and
vein masks \citep{ward_vein_2017} workflows to data collected from a
healthy subject.

@@

While basic Study classes are typically associated with single image
modality or contrast, the analysis suites implemented by them can be
integrated into multi-modal analysis by aggregating multiple Study
classes (sub-studies) in a \emph{MultiStudy} class. Analysis suites are
integrated by joining the data specification tables of the sub-studies
of a MultiStudy class. Entries in specification tables of sub-studies
are joined by mapping them to a common entry in the specification table
of the MultiStudy (Figure \ref{fig:example_multi_study}). This enables derivatives from one
sub-study (e.g. brain extracted T1-weighted anatomical) to be referenced
by workflows of other sub-studies (e.g. anatomically constrained DWI
tractography).

@@

The DwiStudy class implements the extraction of diffusion tensor metrics,
FA and ADC, as well as whole-brain tractography using streamlines tracking from the MRtrix
toolbox \citep{tournier_improved_2010,tournier_mrtrix:_2012} (Figure \ref{fig:dwi_study}).

\begin{figure*}
	\centering
    \includegraphics[width=0.55\textwidth]{../figures/dwi_study}
  \caption{Example diffusion-weighted MRI (DWI) study, which can derive
tensor metrics, fractional anisotropy (FA) and apparent diffusion
coefficient (ADC) as well as streamlines fibre tracking. Blue boxes:
acquired (input) data (filesets or fields). Green boxes: derivatives.
Orange ovals: pipelines. Blue and green arrows: acquired and derived
inputs to pipelines, respectively. Orange arrows: outputs of pipelines.
The DWI magnitude image is preprocessed for motion correction and EPI
distortions masked and bias corrected. From the bias corrected image two
branches of analysis can be performed using the same intermediate
derivatives: FA and ADC and/or streamlines fibre tracking.}
\label{fig:dwi_study}
\end{figure*}

@@

Figure \ref{fig:fa_adc}  displays the FA and ADC maps derived from the diffusion MRI
acquisition. The FA map shows high intensity in known white matter
tracts and low intensities in known grey matter regions. The ADC map
shows high intensity in cortical spinal fluid and low intensity through
the rest of the brain.
Figure \ref{fig:tractography} displays the global tractography derived from the DWI
acquisition. The streamlines follow well known white matter tracts such
as the cortico-spinal, fasciculus and corpus callosum. Intermediate
derivatives derived for the FA and ADC analysis, including the
preprocessed and bias-corrected DWI image and a whole brain mask, were
reused in the generation of the streamlines.

@@

MR contrast-specific analyses are implemented in Banana via a chain
of successively specialised Study sub-classes (e.g.
MRI\textgreater{}EPI\textgreater{}DWI) to enable generic processing
steps (e.g. registration) to be shared between classes. While not
necessary, it is recommended to create a subclass specific to the
research study in question and aggregate all related analysis within it,
since such classes can be applied to alternate datasets in order to
reproduce the exact analysis. The \emph{ArcanaPaper} class
(Supplementary material), which contains methods to generate all figures
in the Results section of this manuscript, is an example of this approach.

@@

Arcana's management of intermediate derivatives and provenance
guarantees that the same analysis is applied across the dataset without
necessarily requiring a complete rerun of the analysis. This guarantee
makes it feasible to process data as it is acquired over the course of
long studies, and therefore help detect any problems that might arise
with the acquisition protocol when they occur. In addition, by reusing
shared intermediate derivatives between analyses, such as the
preprocessed DWI shared between tensor and fibre tracking workflows
(Figure \ref{fig:fa_adc} and \ref{fig:tractography}), processing time as
well as time required for manual
QC is minimised. Given analyses implemented in Arcana are also able to
be processed on HPC clusters, they scale well to large studies.
